An FPGA Implementation of a Long Short-Term Memory Neural Network

### 2016-10-12 paper/paper.tex
Conference Ready

\documentclass{IEEEtran}

%% Additional Packages
\usepackage{cite}
\usepackage{graphicx}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{siunitx}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
%% Macros
\newcommand{\mb}[1]{\mathbf{#1}}
\renewcommand\IEEEkeywordsname{Keywords}

@@

%% COMMENT THIS FOR SUBMISSION
%%\IEEEauthorblockA{Faculty of Engineering University of Porto \\
%%Porto, Portugal}}
%%\IEEEspecialpapernotice{(Working Version)}
\author{
    \IEEEauthorblockN{Joao Canas Ferreira\IEEEauthorrefmark{1}, Jose Fonseca\IEEEauthorrefmark{2}}\\
    \IEEEauthorblockA{\IEEEauthorrefmark{1}INESC TEC and Faculty of Engineering of the University of Porto, Portugal, jcf@fe.up.pt}\\
    \IEEEauthorblockA{\IEEEauthorrefmark{2}CERN, Switzerland, jose.pedro.castro.fonseca@cern.ch}
}

@@

A single LSTM neuron is presented in Figure~\ref{fig:lstmneuron}. As can be seen in the picture,
we still have the recurrent connections from the regular RNNs, but now there are multiple entry points
that control the flow of information through the network. Furthermore, all the
gates are biased, as defined in Equations~\ref{eq:equationsLSTM}. The role and relevance of the main components can
be summarized as follows.

\begin{figure}[!t]
	\centering
	\includegraphics[width=1.1\linewidth]{figures/lstmneuron.png}
    \caption{A complete LSTM neuron, with all the features as described in~\cite{Graves05}. Source:~\cite{Greff15}}
	\label{fig:lstmneuron}
\end{figure}

@@

It is clear that with increasing $K_G$, the maximum clock speed decreases, and that decrease is steeper for larger values
of $K_G$. This means that there is a critical path in the Matrix-Vector multiplication unit, whose multiplexer becomes
increasingly complex for higher values of $K_G$. On the other hand, when $K_G$ is the same, smaller networks are faster
than larger networks. The fastest design is an $N=4$ and $K_G = 2$ network, with a clock frequency of $158.228$ MHz, and
the slowest one is an $N=32$ and $K_G=4$ network, clocked at $101.523$ MHz. The reference design used for validation in
Section~\ref{sec:res-val} is an $N=8$ and $K_G=2$ network clocked at $154.321$ MHz, which yields a clock period of
\SI{6.48}{\nano\second}. The maximum clock achievable for the VC707 networks is reported in Table~\ref{tab:maxfreq-virtx7}

@@

\begin{table}
	\caption{Maximum clock frequencies for the VC707}
	\label{tab:maxfreq-virtx7}
    \centering
  \begin{tabular}{ | l | c | }
    \hline
     & $Frequency$  \\
    \hline
    $N=64$ & 140.854 MHz \\
    \hline
    $N=128$ & 140.854 MHz\\
    \hline
  \end{tabular}

@@

For the networks synthesized on the VC707, the power consumption figures are reported in Table~\ref{tab:power-virtx7}. Since the device is bigger,
and since the network size is larger, the power consumption is higher than before, but still within reasonable levels given
the complexity of the system.

\begin{table}
	\caption{Power Consumption estimates for several Network Sizes $N$ in the VC707}
	\label{tab:power-virtx7}
    \centering
  \begin{tabular}{ | l | c | }
    \hline
     & $Power$  \\
    \hline
    $N=64$ & 1.34 W \\
    \hline
    $N=128$ & 1.509 W\\
\hline
  \end{tabular}

All these estimates were provided by Vivado, and have a \textbf{medium} confidence level.

>>>>

\documentclass[conference]{IEEEtran}

%% Additional Packages
\usepackage{cite}
\usepackage{graphicx}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{siunitx}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{eso-pic}

%% Macros
\newcommand{\mb}[1]{\mathbf{#1}}
\renewcommand\IEEEkeywordsname{Keywords}

@@

%% COMMENT THIS FOR SUBMISSION
%%\IEEEauthorblockA{Faculty of Engineering University of Porto \\
%%Porto, Portugal}}
%%\IEEEspecialpapernotice{(Working Version)}
\author{
    \IEEEauthorblockN{Joao Canas Ferreira\IEEEauthorrefmark{1}, Jose Fonseca\IEEEauthorrefmark{2}}\\
    \IEEEauthorblockA{\IEEEauthorrefmark{1}INESC TEC and Faculty of Engineering of the University of Porto, Portugal, jcf@fe.up.pt}\\
    \IEEEauthorblockA{\IEEEauthorrefmark{2}Faculty of Engineering of the University of Porto, Portugal, jose.pedro.castro.fonseca@cern.ch}
}

\AddToShipoutPicture*{\small\sffamily\raisebox{1.8cm}{\hspace{1.8cm}978-1-5090-3707-0/16/\$31.00 \copyright2016 European Union}}

@@

The structure of an LSTM neuron is presented in~\cite{Greff15}. LSTM Networks retain recurrent connections from the regular RNNs, but now there are multiple entry points
that control the flow of information through the network. Furthermore, all the
gates are biased, as defined in Equations~\ref{eq:equationsLSTM}. The role and relevance of the main components can
be summarized as follows.

@@

It is clear that with increasing $K_G$, the maximum clock speed decreases, and that decrease is steeper for larger values
of $K_G$. This means that there is a critical path in the Matrix-Vector multiplication unit, whose multiplexer becomes
increasingly complex for higher values of $K_G$. On the other hand, when $K_G$ is the same, smaller networks are faster
than larger networks. The fastest design is an $N=4$ and $K_G = 2$ network, with a clock frequency of $158.228$ MHz, and
the slowest one is an $N=32$ and $K_G=4$ network, clocked at $101.523$ MHz. The reference design used for validation in
Section~\ref{sec:res-val} is an $N=8$ and $K_G=2$ network clocked at $154.321$ MHz, which yields a clock period of
\SI{6.48}{\nano\second}. The maximum clock achievable for the VC707 networks is 140.854 MHz, for $N=64$ and $N=128$.

@@

@@

For the networks synthesized on the VC707, the power consumption was 1.34 Watt for $N=64$, and 1.51 Watt for $N=128$. Since the network size is larger, the power consumption is higher than before, but still within reasonable levels given the complexity of the system. These estimates were provided by Vivado, and have a \textbf{medium} confidence level.
